Estimating initial state and quark-gluon plasma medium properties \\ using a hybrid model with nucleon substructure \\ calibrated to p-Pb and Pb-Pb collisions at \texorpdfstring{$\mathbf{\sqrts=5.02}$}{}~TeV

### 2018-07-25 paper/substructure.tex
More plucking away at MAP results and conclusion section

A simple remedy is to replace the arithmetic mean of the wounded nucleon model with a more flexible parametrization
\begin{equation}
  \label{eq:gmean}
  \frac{dS}{dy}\bigg\vert_{y=0} \propto \left( \frac{\T_A^p + \T_B^p}{2} \right)^{1/p},
\end{equation}
based on a family of functions known as the generalized mean(s).
This parametrization introduces a dimensionless parameter $p$ which varies the scaling behavior of initial entropy deposition at midrapidity.
For certain discrete values of $p$, it reduces to well known functional forms such as the arithmetic, geometric and harmonic means:
\begin{equation}
  \newlength{\extraspace}
  \setlength{\extraspace}{0.5ex}
  \frac{dS}{dy}\bigg\vert_{y=0} \propto
  \begin{cases}
    \max(T_A, T_B) & p \rightarrow +\infty, \\[\extraspace]
    (T_A + T_B)/2 & p = +1, \hfill \text{ (arithmetic)} \\[\extraspace]
    \sqrt{T_A T_B} & p = 0, \hfill \text{ (geometric)} \\[\extraspace]
    2\, T_A T_B/(T_A + T_B) & p = -1, \hfill \text{ (harmonic)} \\[\extraspace]
    \min(T_A, T_B) & p \rightarrow -\infty.
  \end{cases}
  \label{eq:gmean}
\end{equation}
Conveniently, it has been shown that the generalized mean ansatz is able to mimic the scaling behavior of certain saturation based initial condition models \cite{Bernhard:2016tnd}, and hence it should serve as a reasonable parametric form for exploring QGP entropy deposition, assuming imperfect knowledge of saturation effects in nature.
Of course the model is \emph{not} a substitute for first principle theory calculations, and it may fail to reproduce nuanced features of \emph{ab initio} models such as the existence of short-range gluon field fluctuations \cite{Schenke:2012wb}.

@@

\subsection{Initial condition parameters}

@@

The \trento\ normalization factor ${\text{Norm} = 20.3^{+2.2}_{-2.2}}$ and generalized mean parameter $p=0.018^{+0.146}_{-0.159}$ are well constrained by the present analysis.
Moreover, their posterior values nicely describe the $p$-Pb and Pb-Pb calibration observables in Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}.
While it would not be surprising, for example, to fit one or two of these observables using such a narrow range of values, the quality of the combined fit and the number of observables described is highly non-trivial.
For example, consider the ratio of the $p$-Pb yield over the Pb-Pb yield, which imposes a strong constraint on physically reasonable initial condition models.
As the entropy deposition parameter $p$ trends toward positive (negative) infinity, particle production scales like the local maximum (minimum) of the nuclear overlap density.
The parameter $p$ thus strongly affects the $p$-Pb and Pb-Pb yield ratio.
It just so happens that the small range of $p$-values needed to describe this yield ratio are the same values needed to describe all of the calibration observables in the present study, \emph{and} numerous Pb-Pb observables at $\sqrts=2.76$~TeV \note.
This work thus reaffirms an empirical scaling law reported in several previous studies \note, that entropy deposition at midrapidity in ultrarelativistic nuclear collisions scales approximately as
\begin{equation}
  \label{eq:geometric_mean}
  \frac{dS}{dy}\bigg\vert_{y=0} \propto \sqrt{\T_A \T_B},
\end{equation}
where $\T_A$ and $\T_B$ are the participant thickness functions \eqref{eq:fluctuated_thick} of each nucleus.
We emphasize that this specific analytic form should not be interpreted too literally.
For instance, a generalized mean described by $p=0.05$ is well within our 90\% credible interval, but it is not equal to the geometric mean in Eqn.~\ref{eq:geometric_mean}.
This approximate form nevertheless has been shown to mimic the scaling behavior of several saturation-based theory calculations in high-energy QCD \cite{Bernhard:2016tnd}, and thus appears consistent with theoretical expectations.

@@

Meanwhile, the posterior on the constituent number $n_c$ is not sharply peaked.
We therefore refrain from quoting a distribution median and 90\% credible interval, although we do note that the distribution clearly favors $n_c > 1$ constituents.
This is not surprising.
The \trento\ model mimics saturation-based initial condition models, and saturation models tend to produce ``proton-sized'' fireballs in $p$-Pb collisions \note.
When the proton is spherically symmetric, the resulting proton-sized QGP is also largely symmetric and thus produces very little flow.
Saturation-based models are therefore unable to describe the significant flow measured in high-multiplicity $p$-Pb collisions without nucleon substructure, or alternatively, some other source of additional correlations \note.

@@

The full posterior in Fig.~\ref{fig:posterior} is the distribution of true parameter values, given the assumptions of our model and the evidence provided by the experimental data.
Correspondingly, the probability of nature ``choosing'' any specific set of parameter values is \emph{zero}; we can only assign a probability to a finite region of parameter space.
Nevertheless, we can define the \emph{maximum a posteriori probability} (MAP) estimate as the single most probable neighborhood (i.e.\ point) in parameter space, equal to the mode of the distribution.

We list the posterior MAP parameters in Tbl.~\ref{tab:map_param}.
The values are estimated using a m

and run \order{6} events minimum-bias and multiplicity triggered $p$-Pb and Pb-Pb events at $\sqrts=5.02$~TeV.
We then calculate all of the model observables listed in Sec.~\ref{sec:observables}.
These ``best fit'' model predictions are plotted in Fig.~\ref{fig:obs_map} alongside the experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
They are explicit model calculations; \emph{there is no emulator}.

\begin{table}[t]
  \caption{Maximum a posteriori (MAP) parameters.}
  \begin{ruledtabular}
  \begin{tabular}{lll}
    Parameter         & Description                        & Value\\
    \paddedhline
    Norm              & Normalization factor                 & 20.            \\
    $p$               & Entropy deposition parameter         & 0.0            \\
    $k$               & Constituent gamma fluct.\ shape      & 0.19           \\
    $w$               & Nucleon width parameter              & 0.92 fm        \\
    $n_c$             & Number of nucleon constituents       & 6              \\
    $v$               & Constituent width                    & 0.43 fm        \\
    $\dmin$           & Minimum inter-nucleon distance       & 0.81 fm        \\
    $\taufs$          & Free-streaming time                  & 0.37 fm/$c$    \\
    $\eta/s$ min      & Minimum value of $\eta/s$ (at $T_c$) & 0.11           \\
    $\eta/s$ slope    & Slope of $\eta/s$ above $T_c$        & 1.6 GeV$^{-1}$ \\
    $\eta/s$ crv      & Curvature of $\eta/s$ above $T_c$    & -0.29          \\
    $\zeta/s$ max     & Maximum value of $\zeta/s$           & 0.032          \\
    $\zeta/s$ width   & Width of $\zeta/s$ peak              & 0.024 GeV      \\
    $\zeta/s$ $T_0$   & Temperature of $\zeta/s$ maximum     & 175 MeV        \\
    $T_\text{switch}$ & Switching/particlization temp.       & 151 MeV        \\
  \end{tabular}
  \end{ruledtabular}
  \label{tab:map_param}
\end{table}

@@

>>>>

A simple remedy is to replace the arithmetic mean of the wounded nucleon model with a more flexible parametrization
\begin{equation}
  \label{eq:gmean}
  \frac{dS}{dy}\bigg\vert_{y=0} \propto \left( \frac{\T_A^p + \T_B^p}{2} \right)^{1/p},
\end{equation}
based on a family of functions known as the generalized mean(s).
This parametrization introduces a dimensionless parameter $p$ which varies the scaling behavior of initial entropy deposition at midrapidity.
For certain discrete values of $p$, it reduces to well known functional forms such as the arithmetic, geometric and harmonic means:
\begin{equation}
  \newlength{\extraspace}
  \setlength{\extraspace}{0.5ex}
  \frac{dS}{dy}\bigg\vert_{y=0} \propto
  \begin{cases}
    \max(T_A, T_B) & p \rightarrow +\infty, \\[\extraspace]
    (T_A + T_B)/2 & p = +1, \hfill \text{ (arithmetic)} \\[\extraspace]
    \sqrt{T_A T_B} & p = 0, \hfill \text{ (geometric)} \\[\extraspace]
    2\, T_A T_B/(T_A + T_B) & p = -1, \hfill \text{ (harmonic)} \\[\extraspace]
    \min(T_A, T_B) & p \rightarrow -\infty.
  \end{cases}
  \label{eq:trento_p}
\end{equation}
Conveniently, it has been shown that the generalized mean ansatz is able to mimic the scaling behavior of certain saturation based initial condition models \cite{Bernhard:2016tnd}, and hence it should serve as a reasonable parametric form for exploring QGP entropy deposition, assuming imperfect knowledge of saturation effects in nature.
Of course the model is \emph{not} a substitute for first principle theory calculations, and it may fail to reproduce nuanced features of \emph{ab initio} models such as the existence of short-range gluon field fluctuations \cite{Schenke:2012wb}.

@@

\subsection{Initial condition properties}

@@

The \trento\ normalization factor ${\text{Norm} = 20.3^{+2.2}_{-2.2}}$ and generalized mean parameter $p=0.018^{+0.146}_{-0.159}$ are well constrained by the present analysis.
Moreover, their posterior values nicely describe the $p$-Pb and Pb-Pb calibration observables in Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}.
While it would not be surprising, for example, to fit one or two of these observables using such a narrow range of values, the quality of the combined fit and the number of observables described is highly non-trivial.
For example, consider the ratio of the $p$-Pb yield over the Pb-Pb yield, which imposes a strong constraint on physically reasonable initial condition models.
As the entropy deposition parameter $p$ trends toward positive (negative) infinity, particle production scales like the local maximum (minimum) of the nuclear overlap density.
The parameter $p$ thus strongly affects the $p$-Pb and Pb-Pb yield ratio.
It just so happens that the small range of $p$-values needed to describe this yield ratio are the same values needed to describe all of the calibration observables in the present study, \emph{and} numerous Pb-Pb observables at $\sqrts=2.76$~TeV \cite{Bernhard:2018hnz}.
This work thus reaffirms an empirical scaling law reported in several previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz}, that entropy deposition at midrapidity in ultrarelativistic nuclear collisions scales approximately as
\begin{equation}
  \label{eq:geometric_mean}
  \frac{dS}{dy}\bigg\vert_{y=0} \propto \sqrt{\T_A \T_B},
\end{equation}
where $\T_A$ and $\T_B$ are the participant thickness functions \eqref{eq:fluctuated_thick} of each nucleus.
We emphasize that this specific analytic form should not be interpreted too literally.
For instance, a generalized mean described by $p=0.05$ is well within our 90\% credible interval, but it is not equal to the geometric mean in Eqn.~\ref{eq:geometric_mean}.
This approximate form nevertheless has been shown to mimic the scaling behavior of several saturation-based theory calculations in high-energy QCD \cite{Bernhard:2016tnd}, and thus appears consistent with theoretical expectations.

@@

Meanwhile, our posterior on the constituent number $n_c$ is not sharply peaked.
We therefore refrain from quoting a distribution median and 90\% credible interval, although we do note that the distribution clearly favors $n_c > 1$ constituents.
This is not surprising.
The \trento\ model mimics saturation-based initial condition models \cite{Bernhard:2016tnd}, and saturation models tend to produce ``proton-sized'' fireballs in $p$-Pb collisions \cite{Bzdak:2013zma}.
When the proton is spherically symmetric, the resulting proton-sized QGP is also largely symmetric and thus produces very little flow.
Saturation-based models are therefore unable to describe the significant flow measured in high-multiplicity $p$-Pb collisions without nucleon substructure, or alternatively, some other source of additional correlations \cite{Schenke:2017bog}.

@@

In this section we select a single set of high-probability model parameters

We verify the emulator predictions and model accuracy by running a large number of $p$-Pb and Pb-Pb events using a single set of high-probability parameters selected from the mode of the posterior distribution.
These parameters approximate the ``best fit'' values
We then calculate all of the model observables listed in Sec.~\ref{sec:observables}.
%These ``best fit'' model predictions are plotted in Fig.~\ref{fig:obs_map} alongside the experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
%They are explicit model calculations; \emph{there is no emulator}.
%

\begin{table}[t]
  \caption{
    \label{tab:mode_params}
    High-probability parameters selected from the posterior distribution and used to generate Fig.~\ref{fig:obs_map}.
    The posterior distribution on the particlization temperature $T_\text{switch}$ is flat (agnostic), so we fix it's value using Ref.~\cite{Bernhard:2018hnz}.
  }
  \begin{ruledtabular}
    \begin{tabular}{ll@{\hspace{2em}}ll}
      \multicolumn{2}{c}{Initial condition} & \multicolumn{2}{c}{QGP medium} \\
      \paddedhline
      Norm     & 20.            & $\eta/s$ min      & 0.11           \\
      $p$      & 0.0            & $\eta/s$ slope    & 1.6 GeV$^{-1}$ \\
      $k$      & 0.19           & $\eta/s$ curv     & -0.29          \\
      $n_c$    & 6              & $\zeta/s$ max     & 0.032          \\
      $w$      & 0.92 fm        & $\zeta/s$ width   & 0.024 GeV      \\
      $v$      & 0.43 fm        & $\zeta/s$ $T_0$   & 175 MeV        \\
      $\dmin$  & 0.81 fm        & $T_\text{switch}$ & 151 MeV        \\
      $\taufs$ & 0.37 fm/$c$
    \end{tabular}
  \end{ruledtabular}
\end{table}

@@

In this work, we constructed a multi-stage hybrid transport model for relativistic nuclear collisions and calibrated free parameters of the model using Bayesian parameter estimation.
Specifically, we focused on improving the model's description of small collision systems by investigating the effect of nucleon substructure on bulk observables in $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV.

The analysis was performed using a parametric data-driven approach to account for theoretical uncertainty (variability) in the QGP initial conditions and hydrodynamic transport coefficients using a handful of equally uncertain model parametersi.
We marginalized over the theoretical uncertainty in the parameter using a handful of equally uncertain model parameters to vary the size and shape of the nucleon, the scaling behavior of initial QGP entropy deposition, the duration of pre-equilibrium dynamics, and the temperature dependence of the hydrodynamic shear and bulk viscosities.

We placed a conservative prior on the ``true'' value of each parameter, and calculated a global likelihood

varied each parameter within its design range to explored the full 15-dimensional parameter space of our model.

constructed an emulator to interpolate the predictions of our model at arbitrary regions of parameter space.

Bayesian inference was then used to compare the predictions of the model to experimental data from CMS and ALICE, in order to quantify the posterior probability that a given set of model parameters describes the experimental data
